Network Deliberation: The role of network structure in large-scale, internet-enabled, participatory decision-making

### 2022-05-05 UMich/chapters/ch04-NetDelibExp.tex
Import NetDelibExp and formatting fixes.

\section{Introduction}
Networks of small interlocking pods appear as a common thread throughout
successful large-scale collaborations.
However, it remains unclear whether that network structure plays a causal role
in the success of those collaborations,
and whether the specific structural properties of that interlocking network matter.
Observational study suggests a correlation between properties such as degree,
structural efficiency, and structural inequality and the performance/productivity
of collaborations \cite{platt_network_2018}.
Numerical simulation suggests a causal relationship in a simplified theoretical
setting.
This chapter proposes an online experiment to bridge these two findings.

\section{Study Software}

In order to conduct studies of network deliberation online, I have modified the
free and open-source Loomio platform \cite{jackson_open_2016}
to implement network deliberation.
Loomio is a widely-used online deliberation platform which provides both
forum functionality (Figure \ref{fig:reply})
and various forms of voting, including ranked-choice (Figure \ref{fig:vote}).
Loomio is popular among grassroots organizations and worker-owned cooperatives
(Loomio itself is a worker-owned cooperative).
Loomio was chosen for its combination of built-in functionality,
existing user base, and open-source extensibility.

The modified platform implements both random-pod and long-path assignment methods
and automatically reassigns participants when the stage is advanced.
In addition to network deliberation features, I have made several modifications to implement
the experimental protocol.
Upon signing in for the first time,
participants are randomly assigned to a treatment group.
Participants are also assigned an alias which they will use throughout the
deliberation (Figure \ref{fig:alias}).

\begin{figure}
\label{fig:reply}
\center
\frame{\includegraphics[width=5in]{fig/Experiment/fig-reply.png}}
\caption{Participants can perform standard forum actions, such as posting and commenting.}
\end{figure}

\begin{figure}
\label{fig:vote}
\center
\frame{\includegraphics[width=3in]{fig/Experiment/fig-vote.png}}
\caption{Participants are presented with a ranked-choice voting screen between rounds of deliberation.}
\end{figure}

\begin{figure}
\label{fig:alias}
\center
\frame{\includegraphics[width=3in]{fig/Experiment/fig-alias.png}}
\caption{Participants are assigned an alias for the duration of the experiment.}
\end{figure}

\section{Pilot Studies}
Several pilot studies have already been conducted:
one in a synchronous lab setting, and two using the asynchronous online
software.
In addition to providing an opportunity to test software,
these pilot studies have provided crucial information for developing a
feasible experimental design.

The initial pilot study was conducted on December 21, 2018,
before the custom online platform was developed.
In this pilot, 18 participants engaged in a synchronous 60-minute deliberation.
Participants sat at individual computer workstations, separated by dividers.
Participants discussed the following question:
``Where should the use of electric scooters be permitted?''
Participants were placed in a single treatment group, and divided into 4
pods of size 4--5.
These pods were reassigned using the random-pod method for a total of three
rounds.
The deliberation took place in chat rooms running the free and open-source
OTree online experiment software \cite{chen_otreeopen-source_2016}.

The second pilot took place after the initial development of the custom online
platform, from August 24, 2020 through August 32, 2020.
In this pilot, 7 participants engaged in an asynchronous deliberation over the
course of 7 days.
Participants were assigned to a single treatment group using network deliberation
with random-pod assignment.
Pods were size 4, and there were 3 rounds (2, 2, and 3 days).
Participants were researchers considering a week-long collaboration and
deliberated on the following question:
``What is your current preference for week of science topics?''

The third pilot took place from August 23, 2021 through August 29, 2021.
Participants deliberated on the following question:
``How should schools in the US reopen for fall 2021?''
Participants were recruited on reddit and Twitter.
18 participants were enrolled (excluding several identified as fake accounts).
These participants were divided between a conventional deliberation control
group and a random-pod treatment group.
The random-pod treatment group had two pods of size 5.
Deliberation was divided into 3 rounds (2, 2, and 3 days).

In these pilots, participants showed quantitative and qualitative evidence of
changing their preferences over the course of deliberation.
Here is an example of one such interaction:
\begin{quote}
{\bfseries Aspen Sunflower:} Even though my \#1 choice includes mandatory vaccinations, I know it's not possible due to school starting now AND MDHS can't mandate it yet. Based on previous stage, I think most seemed to want vaccinations but also wanted options. One option doesn't fit all students.

{\bfseries Aspen Poppy:} Major gap in my understanding here! The schools do not mandate vaccines for students, the state does. For the 12-17 age group, the vaccine is authorized under emergency use, full FDA approval is pending. This strengthens my desire to see hybrid learning with smaller in person groups and vaccinated staff.
\end{quote}

While the pilots did not have enough participants to identify any differences
in the evolution of preferences between different treatment groups,
evidence that preferences do change through deliberation affirms the central
assumption of this dissertation and further motivates the question of how
communication network structure influences that evolution.

\section{Experimental Design}

This section describes the experimental design of the proposed field experiment.
In this experiment, participants will use a pseudonymous online platform
to deliberate on a policy issue.
The deliberation will take place asynchronously, over a period of several days.
At several times throughout the deliberation,
participants will complete a ranked-choice poll regarding the policy issue,
allowing the evolution of their preferences to be tracked over the course of the
experiment.
Depending on opportunity and necessity, multiple waves of deliberation may
be conducted, each with a distinct policy issue and set of participants.

\subsection{Deliberative Topic}
The policy issue chosen as the topic of deliberation is a crucial component
of the experiment.
The topic will be chosen carefully according to several criteria.
The chosen policy issue must be:
\begin{itemize}
\item Relevant to the experimental population;
\item Amenable to participants changing their preferences based on new information or reasoning;
\item Amenable to a predefined list of proposed solutions;
\item Sufficiently complex to have three or more proposed solutions;
\item Timely, but unlikely to be influenced by current events during the deliberation.
\end{itemize}
Examples of policy issues meeting these criteria include:
\begin{itemize}
\item ``Where should the use of electric scooters be permitted?''
\item ``Given the COVID-19 pandemic, how should schools in the US reopen in Fall 2021?''
\end{itemize}

\subsection{Participant Recruitment and Incentives}
Participants will be recruited using several methods.
Participants will be recruited individually through posts on web forums
relevant to the policy question (e.g., reddit, facebook groups)
and to relevant email lists.
Participant pools will also be utilized,
including lists of previous participants
who have expressed interest in future studies
and the UMSI Online Recruitment System for Economic Experiments.
Individual participants will also be recruited using promoted posts and
advertisements on Twitter, Facebook, etc.
Separately, participants may also be recruited through partnership with
community groups who wish to try network deliberation as part of an organizational
decision-making strategy.
The number of such collaborations will depend on the number of interested
partner organizations that have decisions amenable to network deliberation
over the course of the experiment.

In order to incentivize participation and to compensate participants fairly
for their time,
participants who complete the deliberation and pre/post-experiment surveys
will be compensated financially.
Participants will be paid a predetermined fixed amount in order to
minimize the effect of the compensation on their deliberative behavior and
expressed preferences.

\subsection{Experimental Protocol}

Before the deliberation period begins,
participants will complete a short demographic survey.
This survey will enable the identification of sample bias in age, gender, race,
and other demographic factors.

Participants will be randomly assigned to one of three treatment/control groups
at time of enrollment:
\begin{description}
\item[Control]{
The control group will engage in conventional single large group deliberation.
Posts and comments made by participants in the control group will be visible
to all others in the control group for the entire duration of the deliberation.}
\item[Network Deliberation (Efficient)]{
Participants in this treatment group will be divided into small
(4--5 participant) pods.
Posts and comments made by participants in this treatment group will only be
visible to others in their current pod.
All participants will be assigned to a new pod at the beginning of each round
of deliberation using the random-pod assignment method,
producing a structurally efficient communication network.}
\item[Network Deliberation (Inefficient)]{
Participants in this treatment group will be divided into small
(4--5 participant) pods.
Posts and comments made by participants in this treatment group will only be
visible to others in their current pod.
All participants will be assigned to a new pod at the beginning of each round
of deliberation using the long-path assignment method,
producing a structurally inefficient communication network.}
\end{description}

The deliberation period will be divided into a predetermined number of {\em rounds},
each lasting a fixed length of time.
Present plans are to divide the deliberation period into three stages of two days
each.
However, these numbers may be adjusted to ensure time timeline of the
deliberation is consistent with any constraints imposed by specific policy issues.
Before and after each stage, participants will be asked to rank the proposed
solutions to the policy question according to their individual preference.
During each round, participants will be shown a discussion prompt and will be
able to post a response to that prompt.
Example prompts are shown in Table \ref{tab:prompts}.
Participants will be able to view and comment on the posts of other participants
assigned to the same pod.
Participants will also be able to view posts and comments that were visible to them
in previous rounds, but will no longer be able to interact with or reply to
those posts and comments.

\begin{table}
\center
\label{tab:prompts}
\begin{tabular}{|p{0.3in}|p{3.6in}|}
\hline
Stage & Prompt \\
\hline
1 &
Which proposals do you prefer, and why?
If there are disagreements, try asking questions to understand the perspective
of others and to understand the causes of the disagreement.
\\
\hline
2 & In the previous round of discussion, what opinions and reasoning did you
observe in your group?
How much agreement was there?
Were there any disagreements or conflicts?
If so, what were the sources of conflict, and how might they be resolved?
\\
\hline
3 &
What seem to be the most popular opinions?
Do you agree with them?
Has your opinion changed over the course of the discussion?
\\
\hline
\end{tabular}
\caption{Example prompts for each round of deliberation.}
\end{table}

\section{Quantitative Analysis}

The primary quantitative data will be the results of the ranked-choice
polls conducted before and after each stage of deliberation.
These results represent individual participants' stated preferences at multiple
time points.
Standard tools from social choice theory can be applied to such rank-ordered
preferences \cite{arrow_social_2012, brandt_computational_2012}.
At each time, these preferences can be aggregated into a {\em preference profile}
representing the collective preferences of all participants in a treatment group.
Preference profiles can be used to aggregate individual preferences into a single
ranking via a {\em social welfare function}.
Similarly, preference profiles can be used to select a single winner via a
{\em social choice function}.
These social choice and social welfare functions represent voting methods.
Different voting methods can potentially yield different results for the same
preference profile.
However, the more similar preferences are across individual participants,
the more likely it is that different voting methods will yield the same winner
or social ranking.
For example, when one proposal beats all others in pairwise contests, it is known as the
{\em Condorcet winner}.
Many voting methods yield the Condorcet winner when it exists, but the existence
of a Condorcet winner is not guaranteed.

For the purposes of this dissertation, the preference profiles will be used
to quantify agreement and conflict over the course of deliberation.
Within social choice theory, agreement can be defined by a {\em consensus class},
a set of preference profiles meeting one of several consensus criteria
\cite{elkind_rationalizations_2016}.
Examples include strong unanimity (of rank orders),
unanimity (of winners),
majority (existence of majority),
Condorcet (existence of Condorcet winner),
and transitivity (of social preference).
On the most restrictive end, in strong unanimity and unanimity,
all decision-makers prefer the same alternative.
In contrast, transitivity only requires that the social preferences induced by
individual preferences are transitive.
Distance-rationalizable voting systems are defined by projecting the preference
profile onto the nearest consensus profile according to a suitable distance.
For example, Dodgson's method \cite{dodgson_method_1876, brandt_computational_2012}
defines a distance based on swapping adjacent entries in individual preference profiles.
However, for most voting systems,
finding the consensus profile that minimizes distance is NP-Hard
\cite{elkind_rationalizations_2016}.
Social choice theory also provides measures of distances between two individual
preference rankings.
These measures can be extended to create measures of agreement for entire
preference profiles that can be calculated efficiently.

One possibility is to measure the distance from strong unanimity.
Two such methods will be used to analyze the results of the deliberation experiment:
the mean Kendal tau metric and the mean Spearman correlation.
The Kendal tau metric is the number fraction of pairwise contests that have
different results between two profiles \cite{kendall_new_1938}.
By averaging this measure over all pairs of members, an overall measure of agreement can be calculated for the group.
This measure does not incorporate information about win/loss margins,
only whether an alternative wins or loses.
To quantify the win/loss margin, the mean Spearman correlation will be used.
The Spearman correlation is the linear coefficient of correlation between the
rank orders of alternatives for two voters \cite{spearman_proof_1904}.
As with Kendall tau, the Spearman correlation can be averaged over all pairs of
members to determine a measure of agreement for the entire group.
It must be noted that representing agreement with a single number will inevitably
ignore important information.
Neither of these measures distinguishes a small number of voters disagreeing
on many comparisons from a large number of voters disagreeing on a few.
The Spearman correlation is also unable to distinguish these from situations in
which a small number of voters disagree by a very large margin.

While unanimity might be preferred,
it may sometimes be more realistic to focus on transitivity.
This consensus class requires only that the social preferences between pairs
of alternatives are transitive.
Such preference profiles guarantee that a Condorcet winner will exist,
bringing many different voting methods into agreement on which should win.

The deviation from transitivity will be qunatified by counting the number
of ranked-pair violations.
When performing a vote by the Tideman ranked pair system,
the relative order of two alternatives in the social preference is
determined by the winner of a pairwise vote,
except in the special case that contests with a higher margin have already
constrained their order due to transitivity.
In the latter case, the winner of the pairwise contest may have a lower
position in the social preference.
If this situation occurs, the social preference is intransitive and there may
not be a Condorcet winner.
Each such pair represents a contest violating the social ranking,
so the fraction of such pairs
can be used to quantify the strength of intransitivity in the group’s preferences.
This measure has the property that a Condorcet winner is guaranteed when the value is 0.

\section{Power Analysis}
The minimum number of participants required is determined by a number of factors.
First, the long-path pod assignment method results in a minimum number of pods,
determined by the chosen parameters.
Specifically, the pod assignments for each round beyond the first are determined
by a unique prime number,
and the number of pods is a multiple of that prime.
For $T=3$ rounds, two primes are necessary and choosing the lowest two (2, 3)
yields the lowest minimum number of pods: 3.
For all pods to have at least 4 members, the minimum number of participants in
the long-path treatment is 12.

@@

Combining the above considerations,
an absolute minimum of 12 participants per treatment is necessary to conduct
a wave of the experiment,
and a minimum of about 35 participants per treatment group is necessary to
observe differences between the network deliberation treatment conditions.

\begin{figure}
\label{fig:kcon}
\includegraphics[width=5in]{fig/Experiment/fig-kcon.eps}
\caption{The k-connectivity of long-path and random-pod network deliberation
networks for $T=3$ stages and a pod size of $M=5$.}
\end{figure}

\section{Qualitative Analysis}
In addition to quantitative data,
qualitative data will also be collected in conjunction with
the deliberation experiment.
These data will be used to corroborate the results of quantitative
analysis and to achieve a richer understanding of individual participants'
experience of network deliberation,
as well as to as well as identify common themes that may provide
insight into the strengths and weaknesses of network deliberation.

The text of participants' posts and comments throughout the deliberation
will be kept for analysis.
This text can potentially be used to identify deliberative behaviors that
lead to changes in preferences.
Similarly, this text can be used to track the diffusion of specific keywords,
arguments, and influence through the communication network.

A subset of participants will be randomly selected to participate in
60-minute semi-structured interviews about their experience.
The interview script can be found in Appendix \ref{apx:interview}.
These interviewed will be transcribed and qualitatively coded using a
grounded theory approach.
The initial code book will be developed by open-coding pilot study transcripts
and an initial subset of interview transcripts.

>>>>

\section{Introduction}
Networks of small interlocking pods appear as a common thread throughout
successful large-scale collaborations.
However, it remains unclear whether that network structure plays a causal role
in the success of those collaborations,
and whether the specific structural properties of that interlocking network matter.
Observational study suggests a correlation between properties such as degree,
structural efficiency, and structural inequality and the performance/productivity
of collaborations \cite{platt_network_2018}.
Numerical simulation suggests a causal relationship in a simplified theoretical
setting (Chapter \ref{chap:abm}).
To bridge these two findings,
this chapter presents the results of an online experiment studying the role of network deliberation in the evolution of group preferences.

Collaboration and collective action allow groups of individuals to manage shared resources and combine efforts towards common goals, but also introduce the problem of governance and collective decision-making.
While collective governance of common resources is a historically hard problem, Ostrom described principles seen in successful systems \cite{ostrom_collective_2000}.
More recently, the proliferation of inexpensive point-to-point global communication via the internet has enabled a number of successful large-scale collaborations and movements, including free/open-source software \cite{benkler_coases_2002, coleman_coding_2012, raymond_cathedral_1999}, Wikipedia \cite{keegan_evolution_2017, benkler_coases_2002, giles_internet_2005}, and social movements \cite{tufekci_twitter_2017, gonzalez-bailon_networked_2016}.

Scholars of democracy commonly describe governance systems in terms of two axes: representation and argumentation \cite{ackerman_deliberation_2002, gonzalez-bailon_networked_2016}. Here, {\em representation} refers to how many people participate in decision-making, while {\em argumentation} refers to the amount of discourse between the decision-makers.
While both can improve governance, they are difficult to achieve simultaneously \cite{gentry_consensus_1982, freeman_tyranny_1972} and tradeoffs can be made to favor one over the other, as in voting (high representation, low argumentation) and representative democracy (low representation, high argumentation).

We focus on deliberation due to fundamental limitations of voting.
Results from social choice theory showt that any ranked-choice voting system is subject to limitations such as the Condorcet Paradox \cite{condorcet_essay_1785} and Arrow's Impossibility Theorem \cite{arrow_social_2012}.
Deliberation offers one way to avoid these limitations by changing individual preferences \cite{ackerman_deliberation_2002}.
The success of deliberation at generating consensus has been attributed to its ability to identify and resolve conflict \cite{gentry_consensus_1982}.
However, deliberation also carries risks, such as the posibility of pushing individuals towards extreme and polarized preferences \cite{schkade_what_2007}.

In any sufficiently large group, it becomes unfeasible for all members to interact with each other.
Understanding collaboration in large groups thus requires attention to the network structure of who interacts with whom.
The field of social learning has studied both simple estimation tasks \cite{degroot_reaching_1974, golub_naive_2010} and more complex rugged-landscape optimization tasks \cite{lazer_network_2007, barkoczi_social_2016}.
Observational studies have been used to analyze the network structures of online collaborations \cite{gonzalez-bailon_networked_2016, platt_network_2018}.
Lab and field studies have also been used to study real-world collaborations on a small scale \cite{salehi_hive_2018, kearns_experiments_2012}.
As discussed in Chapter \ref{chap:abm}, the successful large-scale collaborations enabled by the internet often exhibit network sructures characterized by small, interlocking groups, reminiscent of interlocking directorates \cite{levine_study_1979} and interlocking publics \cite{habermas_structural_1991}.
In Chapter \ref{chap:abm}, we describe agent-based simulations showing that such network deliberation can improve performance on complex tasks when individuals exhibit strong social influence.dsd
Here, we expand on existing research by performing a controlled online experimental evaluation of network deliberation.
We use periodic polling to track preference evolution, allowing causal inference in a real-world setting.

While many studies have examined the role of network structure in collaboration, there has so far been little attention to the interlocking pod structure of network deliberation.
The closest we are aware of is the study of network rotation by Salehi \& Bernstein \cite{salehi_hive_2018}.
The present experiment goes beyond existing work primarily by tracking preferences throughout the experiment and by the use of a conventional single-group control condition.
We also describe multiple network deliberation topologies for achieving efficient or inefficient networks.
For the present study, we focus on an efficient topology.

The experiment described here focuses on several issues.
Based on observational and numerical findings that successful large-scale collaborations often exhibit network deliberation struture, we test the following hypothesis:
\begin{description}
\item[H1:] Network deliberation results in higher agreement among participants than single large-group deliberation.
\end{description}
We are also concerned with the dynamics of preference evolution under network deliberation, which may provide insight into possible mechanisms.
One such mechanism proposed in consensus decision-making literature is the ability to identify and resolve conflict.
We thus also seek to address two research questions:
\begin{description}
\item[RQ1:] How do preferences evolve throughout network deliberation?
\item[RQ2:] How effective is network deliberation at identifying and resolving conflict?
\end{description}

We find difference in preference evolution between network deliberation and conventional deliberation. Our primary contributions are:
\begin{itemize}
    \item We find experimental support for the hypothesis that network deliberation is better at facilitating agreement than conventional deliberation.
    \item We find evidence suggesting network deliberation can provide protection against information cascades.
    \item We find no evidence that network deliberation facilitates substantial conflict-resolution, but that it may provide protection against polarization.
\end{itemize}

We present relevant theory in Section \ref{sec:theory}.
We describe the experiment in Section \ref{sec:experiment} and present the results in Section \ref{sec:res}.
We discuss the results in Section \ref{sec:discussion} and conclude in Section \ref{sec:ch04conclusion}.

\section{Theory}
\label{sec:theory}
\subsection{Social Choice Theory}
Our analysis of preferences is foudned in the formalism of social choice theory \cite{arrow_social_2012, gaertner_primer_2009}.
In social choice theory, individual's {\em preferences} are represented by a ranked ordering of {\em alternatives} (e.g., proposals, candidates).
The set of preferences for all individuals in a group is called the {\em preference profile}.
A {\em social welfare function} can be applied to the profile to generate a {\em social preference}: an ordering of alternatives representing the preference of the group as a whole.
A {\em social choice function} represents a voting system: it selects a subset of winners from the available alternatives.

\subsection{Quantifying Agreement}

Social choice theory typically focuses on identifying and describing the limitations of specific voting systems.
Instead, we build on social choice theory to quantify the level of agreement in a group, based in individual profiles.

We use ranked-choice polls to construct preference profiles which are used to quantify agreement and conflict over the course of deliberation.
Within social choice theory, agreement can be defined by a {\em consensus class},
a set of preference profiles meeting one of several consensus criteria
\cite{elkind_rationalizations_2016}.
Examples include strong unanimity (of rank orders),
unanimity (of winners),
majority (existence of majority),
Condorcet (existence of Condorcet winner),
and transitivity (of social preference).
On the most restrictive end, in strong unanimity and unanimity,
all decision-makers prefer the same alternative.
In contrast, transitivity only requires that the social preferences induced by
individual preferences are transitive.
Distance-rationalizable voting systems are defined by projecting the preference
profile onto the nearest consensus profile according to a suitable distance.
For example, Dodgson's method \cite{dodgson_method_1876, brandt_computational_2012}
defines a distance based on swapping adjacent entries in individual preference profiles.
However, for most voting systems,
finding the consensus profile that minimizes distance is NP-Hard
\cite{elkind_rationalizations_2016}.
Social choice theory also provides measures of distances between two individual
preference rankings.
These measures can be extended to create measures of agreement for entire
preference profiles that can be calculated efficiently.

One possibility is to measure the distance from strong unanimity.
Any distance metric $\delta$ with a range of $[0,1]$ can be converted to a measure of correlation with range $[-1, 1]$ by taking $1 - 2\delta$.
One such measure is based on the Kendall tau metric, and the corresponding Kendall tau correlation \cite{kendall_new_1938}.
The Kendall tau metric is the fraction of pairwise contests that have different results between two profiles.
By averaging the corresponding correlation over all pairs of members, an overall measure of agreement can be calculated for the group.
While commonly used in social choice theory, the Kendall tau weights all contests equally, regardless of the ranks of the alteratives.
In practical settings with a small number of winners, the contests between highly-ranked alternatives are more consequential than those of low-ranked alternatives;
participants don't care if they disagree on the ordering of losing alternatives as long as they agree on which ones win and lose.
Various weighted extensions of the Kendall metric have been proposed \cite{shieh_weighted_1998, can_weighted_2014},
each with benefits and drawbacks.
We propose our own based on metric structure, uniqueness, and efficiency of calculation, which we describe below.

\subsubsection{Weighted Crossing Distance}
We propose the following weighted distance metric for preference rankings.
Given a set of $M$ alternatives, there are $M-1$ ways to divide an ordering into two nonempty sets of adjacent alternatives, i.e., a high set and a low set.
For a pair of preferences $p, q$, we define a length $M - 1$ {\em crossing vector} $v_m(p, q)$ to be equal to the number of alternatives in the highest $m$ positions of $p$ and not in the highest $m$ positions of $q$ (it is easily shown that this measure is symmetric in $p$ and $q$).
Informally, $v_m$ is the number of alternatives that cross place $m$.
Each element of $v_m$ is an integer in the range $[0, v_m^{\max}]$ dependent on the index:
\begin{equation}
    v_m^{\max} =
    \begin{cases}
    m & m \leq (M - 1) / 2 \\
    M - m & (M - 1) / 2 < m \leq M - 1.
    \end{cases}
\end{equation}
We construct the weighted crossing distance by taking a weighted sum over this vector.
There are many potential choices for weights.
We choose the weight of place $m$ to be equal to the highest total of all lower ranked places plus one.
Thus the weight at index $i = M - m$ (noting that low rank corresponds to high $m$) is given by:
\begin{equation}
    w_i = 1 + \sum_{k=1}^{i-1} w_k v_k^{\max}.
\end{equation}

\subsection{Quantifying Social Influence}

We are particularly interested in the role of social influence in preference evolution and require a formalism to quantify and analzye social influence.
If two individuals preferences become closer, can we say whether one influenced the other? And if so, can we say which influence which?
We cannot answer either question with certainty from preferences alone, but we can construct a reasonable proxy based on some simplifying assumptions.
Namely, we will construct measures of social influence based on the assumptions that:
\begin{itemize}
    \item The magnitude of an individual's change in preference increases with their susceptibility with social influence;
    \item The magnitude of an individual's change in preference increases with the magnitude of the distance between their preference and the influencing preference.
\end{itemize}
Thus, if the distance between the preferences of individuals $A$ and $B$ decreases, and the preference of $B$ changed more than the preference of $A$, we assume that individual $A$ influenced individual $B$.
We now describe two formal measures that use these assumptions to quantify how susceptible to influence and how potential influential an individual is.

\subsubsection{Conformity and Hipness}

Let $X$ be a set of alternatives (e.g., proposals or candidates)
and let $\mathcal{L}$ be the set of strict total orders over $X$.
We represent the preference of an individual $p$ at time $t$
by a strict total order $R_p(t) \in \mathcal{L}$.
Given $n$ individuals, we call the $n$-tuple of their preferences
at time $t$ the profile $R(t)$.

Let $d:\mathcal{L}\times\mathcal{L} \to \mathbb{R}$ be a
dissimilarity function on preferences.
We define the mean dissimilarity of the profile $R$ at time $t$ as:
\begin{equation}
    \frac{1}{n(n-1)} \sum_p \sum_{q \neq p} d\left(p_t, q_t\right),
\end{equation}
where $p_t$ is shorthand for $R_p(t)$.
The above quantity can be interpreted as a measure of disagreement among a group of individuals.

Let $t_i$ be an initial time, and $t_f$ be some later time.
The negative change in mean dissimilarity is given by:
\begin{equation}
    \label{eqn:Delta}
    \Delta =
    \frac{-1}{n(n-1)} \sum_p \sum_{q \neq p}
    \left[
    d\left(p_f, q_f)\right)
    - d\left(p_i, q_i\right)
    \right].
\end{equation}
This quantity can be interpreted as the increase in agreement within the group between times $t_i$ and $t_f$,
with negative values signifying a decrease in agreement.

We wish to quantify whether a change in agreement/disagreement is due to individuals converging towards initially popular alternatives, or to the the group being persuaded towards initially unpopular alternatives.
In the former case, we would expect the majority of individuals to change their preferences by a small amount, while a smaller number of individuals holding unpopular opinions make larger shifts towards popular preferences.
In the latter case, we would expect the opposite:
some individuals with unpopular preferences hold closely to those opinions, while a large number of individuals move away from the previous popular preferences.
So we are interested not just in the total change in agreement/disagreement,
but also in how that change is distributed across the individuals in the group.

To apportion the change in agreement between individuals $p$ and $q$,
we make the assumption that an individual's contribution to the change in agreement is proportional to the change in their own preference.
Specifically, we define the conformity $c(p,q)$ of individual $p$ toward individual $q$ as:
\begin{eqnarray}
    d_i &=& d(p_i, q_i) \\
    d_f &=& d(p_f, q_f) \\
    |p| &=& d(p_i, p_f) \\
    c(p,q) &=& -(d_f - d_i)\frac{|p|}{|p| + |q|}.
\end{eqnarray}
Note that the change in agreement between two individuals can be written in terms of the conformity:
\begin{equation}
    -(d_f - d_i) = c(p, q) + c(q, p).
\end{equation}
Similarly, the total change in agreement (\ref{eqn:Delta}) can be rewritten as a sum of conformity over all pairs of individuals:
\begin{equation}
    \Delta =
    \sum_p \sum_{q \neq p} c(p, q).
\end{equation}
By performing only one of the sums above, the change in agreement $\Delta$ can be written as a sum of terms, each corresponding to one individual.
The sum can be performed over either the first or second argument of $c$.
If the sum is performed over the second argument of $c$, the result is the mean conformity of a particular individual $p$ toward all others:
\begin{equation}
    C(p) = \sum_q c(p, q).
\end{equation}
Alternatively, if the sum is performed over the first argument of $c$, the result is the mean conformity of all others toward $p$, which we call the hipness $H(p)$:
\begin{equation}
    H(p) = \sum_q c(q, p).
\end{equation}
Both $C(p)$ and $H(p)$ are equal to $\Delta$ when summed over all individuals, but they describe conceptually different attributes of an individual.
The conformity quantifies how much an individual has increased agreement by changing their preferences to match others.
The hipness quantifies how much an individual has increased agreement due to holding an unchanging preference near to others' final preferences.

\subsection{Network Topologies}
In Chapter \ref{chap:abm}, we describe two models of network deliberation structure: random-pod and long-path.
In order to expect a measurable difference from the control condition and between each other, a minimum network size is required.
We decribe these considerations below.
For the present experiment, we opt for only a control and random-pod condition based on the avaialble number of participants.

\subsubsection{Network Size}
The minimum number of participants required is determined by a number of factors.
First, the long-path assignment method results in a minimum number of pods,
determined by the chosen parameters.
Specifically, the pod assignments for each round beyond the first are determined
by a unique prime number,
and the number of pods is a multiple of that prime.
For $T=3$ rounds, two primes are necessary and choosing the lowest two (2, 3)
yields the lowest minimum number of pods: 3.
For all pods to have at least 4 members, the minimum number of participants in
the long-path treatment is 12.

@@

Combining the above considerations,
an absolute minimum of 12 participants per treatment is necessary to conduct
a wave of the experiment,
and a minimum of about 35 participants per treatment group is necessary to
observe differences between the random-pod and long-path conditions.

\begin{figure}
\label{fig:kcon}
\center
\includegraphics[width=5in]{chapters/figures/NetDelibExp/fig-kcon}
\caption{The k-connectivity of long-path and random-pod network deliberation
networks for $T=3$ stages and a pod size of $M=5$.}
\end{figure}

\section{Experimental Study}
\label{sec:experiment}

\subsection{Experimental Design}

Participants used a pseudonymous online platform (described in Section \ref{sec:exp-platform}) to deliberate on a policy issue.

The deliberation took place asynchronously over a period of 7 days, from Monday, November 15, 2021 through Sunday, November 21, 2021.
The deliberation was divided into 3 rounds of 2--3 days each.
During each round, participants were shown a discussion prompt and were able to post a response to that prompt (Table \ref{tab:prompts}).
Participants were able to view and comment on the posts of other participants assigned to the same pod.
Participants were also able to view posts and comments that were visible to them in previous rounds, but could only interact or reply to posts and comments in the current round.
Before and after each round, participants completed a ranked-choice poll regarding the deliberation topic, allowing the evolution of their preferences to be tracked over the course of the experiment.

\begin{table}
\center
\label{tab:prompts}
\begin{tabular}{|p{0.5in}|p{4.5in}|}
\hline
Stage & Prompt \\
\hline
1 &
Which proposals do you prefer, and why?
If there are disagreements, try asking questions to understand the perspective
of others and to understand the causes of the disagreement.
\\
\hline
2 & In the previous round of discussion, what opinions and reasoning did you
observe in your group?
How much agreement was there?
Were there any disagreements or conflicts?
If so, what were the sources of conflict, and how might they be resolved?
\\
\hline
3 &
What seem to be the most popular opinions?
Do you agree with them?
Has your opinion changed over the course of the discussion?
\\
\hline
\end{tabular}
\caption{Prompts shown to participants during each round of deliberation.}
\end{table}

Participants were recruited from students in an undergraduate university course and received extra credit for participating. 65 participants enrolled and completed at least one round of deliberation.
Basic demographic information was collected and is shown in Table \ref{tab:demographic}.

\begin{table}[]
    \centering
    \begin{tabular}{l l c c}
\hline
\multicolumn{2}{l}{Demographic} & Control & Random-Pod\\
\hline
Age & 18--24 & 30 & 31 \\
& 25--29 & 1 & 0 \\
& 30--34 & 0 & 1 \\
& Non-disclosed & 2 & 0 \\
\hline
Gender & Man & 18 & 14 \\
& Non-binary & 0 & 1 \\
& Woman & 13 & 17 \\
& Non-disclosed & 2 & 0 \\
\hline
Race & Asian & 14 & 16 \\
& Black or African American & 1 & 1 \\
& Hispanic & 1 & 0 \\
& White & 13 & 11 \\
& White, Asian & 1 & 4 \\
& Non-disclosed & 3 & 0 \\
\hline
    \end{tabular}
    \caption{Demographic statistics for experiment participants.}
    \label{tab:demographic}
\end{table}

Participants were assigned to one of two conditions at the time of enrollment.
Assignments were uniformly at random.
Stratified sampling would be preferable in order to ensure even demographic distributions, but was not feasible due to software and time constraints.
However, we manually confirmed that each condition had comparable demographics.
The two conditions were:
\begin{description}
\item[Control]{
Participants in the control condition engaged in conventional deliberation in a single large group.
Posts and comments by participants in the control condition were visible
to all others in the control condition for the duration of the deliberation.}
\item[Random-Pod (Network Deliberation)]{
Participants in this condition were divided into small pods ($\leq 5$ participants).
Posts and comments made by participants in this condition were only visible to others in their current pod.
Participants in this condition were assigned to a new pod at the beginning of each round of deliberation using the random-pod assignment method, producing a structurally efficient communication network.}
\end{description}

\subsection{Deliberation Topic}
The policy issue chosen as the topic of deliberation is a crucial component
of the experiment.
The topic was chosen carefully according to several criteria:
\begin{itemize}
\item Relevant to the experimental population;
\item Amenable to participants changing their preferences based on new information or reasoning;
\item Amenable to a predefined list of proposed solutions;
\item Sufficiently complex to have three or more proposed solutions;
\item Timely, but unlikely to be influenced by current events during the deliberation.
\end{itemize}

Participants were students in a university undergraduate course and deliberated on the topic and format of one section of their final exam.
As the outcome of the final poll determined the content of a real portion of their final exam, participants had an intrinsic incentive to advocate for their true preferences.
The question and alternatives presented to participants were:

\begin{displayquote}

The 2021 [redacted] final exam will include a section worth up to 10\% based on the content covered during the first part of the semester. Which of the following options should be chosen for the topic and format of that section of the exam?

\begin{description}
\item[Prop. 1]{Open-ended with partial credit (2 questions, 5 points each) - Ch. 1-2: Intro to networks.}
\item[Prop. 2]{Open-ended with partial credit (2 questions, 5 points each) - Ch. 3: Bridges, clustering, triadic closure, strong and weak ties, and the Strong Triadic Closure Property.}
\item[Prop. 3]{Multiple choice with no partial credit (5 questions, 2 points each) - Ch. 4: Homophily, selection, social influence, and social-affiliation networks.}
\item[Prop. 4]{Multiple choice with no partial credit (5 questions, 2 points each) - Ch. 5: Signed networks and balance theory.}
\item[Prop. 5]{True/false with no partial credit (10 questions, 1 point each) - Ch. 6: Game theory, best responses, and Dominant Strategies, Nash equilibrium.}
\item[Prop. 6]{True/false with no partial credit (10 questions, 1 point each) - Ch. 9:  Auctions.}
\end{description}
\end{displayquote}

\subsection{Experimental Platform}
\label{sec:exp-platform}
